{\renewcommand{\arraystretch}{2}
\begin{longtable}{ | p{6cm} | p{10cm} | }

\hline
Is the question historical and does it have historical significance? & 

Horses played a key role in our pre-industrialized economy.  They were the main source of transportation and were critical to agriculture, so the impact that the Industrial Revolution had on the horse is very historically significant.
\\\hline

Is the answer to the question obvious or a matter of opinion? & 

No.  The question, while it can be speculated, cannot be answered by a simple search, and The Industrial Revolution had a very concrete impact on the horse, it's not a matter of opinion.
\\\hline

What pieces of information will be needed to provide the framework so that you can infer an answer to the question? & 

1) The role horses played before the Industrial Revolution. \n\n

2) What horses did to contribute to the success of the pre-industrialized economy. \n\n

3) The role horses played after the Industrial Revolution. \n\n

4) Specific aspects of the Industrial Revolution that resulted in this change in roles. \newline
\\\hline

Provide at least 3 potential sources of information. & 

1) 2013 article about the role of horses in brewing during the industrial revolution \cite{almer13}. \n\n

2) 1820 image drawn of a horse hauling a dray (low, flat-bed wagon without sides) \cite{schar20}. \n\n

3) 1982 article written about the British Industrial Revolution \cite{musso82}. \n\n

4) 1831 book containing general history and zoological classification of the horse 
\cite{youat31} \newline
\\\hline

\end{longtable}}
