\cvsection{Experience}

\cvevent{Software Engineer Intern, API Platform}{Target Technology}{Jun 2022 -- Aug 2022}{Minneapolis, MN}

\begin{itemize}
    \item Developed a program to dynamically generate reverse proxy configuration.  Used to map incoming traffic to specifed 
      backend clusters.  Written in Go and provides configuration for Envoy Proxy.
    \item Worked in an Agile team of 10, collaborating and communicating with other teams in different disciplines.
\end{itemize}

\cvtag{Go}
\cvtag{Docker}
\cvtag{Kubernetes}
\cvtag{Fastly}
\cvtag{Ruby}
\cvtag{Hugo}
\cvtag{SQL}
\cvtag{Node.js}
\cvtag{JavaScript}
\cvtag{Git}
\cvtag{Jenkins}
\cvtag{Python}

\divider

\cvevent{Undergraduate University Research}{University of Minnesota, Twin Cities}{Jan 2022 -- May 2023}{Minneapolis, MN}

\begin{itemize}
    \item Developed program to automatically generate Parson's style problems for entry level CS students.
    \item Questions are outputed to QTI and GIFT format for use in Canvas and Moodle courses respectively.
    \item Optionally, introduces mutations to input code to provide a variety of question types, with hopes of improving students 
      ability to analyze code.
\end{itemize}

\cvtag{Python}
\cvtag{Express}
\cvtag{Shell Scripting}
\cvtag{XML}
\cvtag{Gift}
\cvtag{QTI}

\divider

\cvevent{Undergraduate Teaching Assistant}{University of Minnesota, Twin Cities}{Jan 2022 -- May 2023}{Minneapolis, MN}

\begin{itemize}
    \item Introduced C++ as well as basic concepts of object oriented programming.
    \item Taught design patterns to speed up development process and SOLID design principles for project maintainability.
    \item Showed how to structure projects with Unified Modeling Language (UML), stressing the importance of program design/analysis.
\end{itemize}

\cvtag{C++}
\cvtag{Design Patterns}
\cvtag{UML}
\cvtag{JavaScript}
\cvtag{HTML}
\cvtag{Docker}

