\documentclass[11pt]{article}

\usepackage{setspace}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{amsfonts} 
\usepackage{mathtools}
\usepackage{relsize}
\usepackage[top=2cm,bottom=2cm,left=2.5cm,right=2.5cm,marginparwidth=1.75cm]{geometry}
\setlength{\parindent}{0cm}

\renewcommand{\l}{\lambda}

\title{4512 Final}
\author{Fletcher Gornick}
\date{December 18, 2021}
\spacing{1}

\begin{document}
 \maketitle 
 \textbf{1) Find the general solution of } \(\mathbf{\displaystyle\frac{d^4 y}{dt^4} + y = 0.}\)
 \newpage

 \textbf{2) We know that the SIR model is given by the equations}
 \[\mathbf{\frac{d}{dt} \begin{pmatrix*}[r] S \\ I \\ R\end{pmatrix*}
   = \begin{pmatrix*}[r] -\beta S I \\ \beta A I - \delta I \\ \delta I\end{pmatrix*}}\]
 \textbf{where the removal rate \(\delta\) and the infection rate \(\beta\) are positive 
 constants. We want to modify this model by changing the assumption that the infected population 
 keeps it's immunity. So we assume that the rate of change of the population susceptible to 
 infection is increased proportionally to \(R\). The new model is}
 \[\mathbf{\frac{d}{dt} \begin{pmatrix*}[r] S \\ I \\ R\end{pmatrix*} = 
 \begin{pmatrix*}[r] -\beta S I \\ \beta A I - \delta I \\ \delta I\end{pmatrix*}} + 
 \begin{pmatrix*}[r] \mu R \\ 0 \\ -\mu R \end{pmatrix*}, \quad \mu > 0.\]
 \textbf{Find the equilibrium points and determine if they are stable or not. Plot the phase 
 portrait. What happens to \((S(t),I(t),R(t))\) as \(t\) goes to infinity? (Consider only 
 \(S,I,R \geq 0\)).}
 \newpage

 \textbf{3) Consider the autonomous system}
 \[\frac{d}{dt} \begin{bmatrix*}[r] x \\ y \end{bmatrix*} = 
   \begin{bmatrix*}[r] x(a-by) \\ y(-c+dx) \end{bmatrix*}, \quad a,b,c,d > 0.\]
 \textbf{Find the equilibrium points (only for \(x,y \geq 0\)) and determine (if possible)
 their stability. Plot the phase portrait. Argue that, if \(x(t=0)\) and \(y(t=0\)) are positive, 
 the corresponding orbit is periodic. (Hint: show that the orbits lie on the curves of the form 
 \(dx -c \ln x + by - a \ln y = \) constant). Plot the phase portrait.}
 \newpage 

 \textbf{4) Use the Poincar\'{e}-Bendixson theorem to prove that a periodic orbit of the system}
 \[\mathbf{\frac{d}{dt} \begin{bmatrix*}[r] x \\ y \end{bmatrix*} = 
   \begin{bmatrix*}[r] \frac{1}{2}x-y - \frac{1}{2}x(x^2+y^2) \\ 
   x+\frac{1}{2}y - \frac{1}{2}y(x^2+y^2) \end{bmatrix*}}\]
 \textbf{lies on the circle \(x^2+y^2=1\). Sketch the phase portrait of the system. What happens 
 to \((x(t),y(t))\) as \(t\) goes to infinity?}
\end{document}
