\documentclass[11pt]{article}

\usepackage{setspace}
\usepackage[top=2cm,bottom=2cm,left=2.5cm,right=2.5cm,marginparwidth=1.75cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[hidelinks,urlcolor=cyan]{hyperref}
\urlstyle{same}

\title{Project Coding Style Research}
\author{Fletcher Gornick}
\date{September 27, 2021}

\spacing{1.5}
\begin{document}
\maketitle

Before giving my novice opinion on the Google C++ Style Guide, I thought I'd take a look at the 
meta surrounding it.  From what I've found, there are very mixed thoughts on Google's approach 
to how C++ code should be written.  It seems as though the public consensus is that this style 
guide is focused most on making code more compatible with Google's existing code base.  That 
matter aside, I do believe that the guide is actually very insightful with only a few minor 
exceptions.  My short answer as to if I'll use Google's C++ style directly, is no, I'll definitely 
use the guide as a sort of template to guide my process, but different projects should be approached 
in different ways.  Although following one format for projects can lead to a very flexible and 
extensible program, it's much better to approach a problem your own way, this is what ultimately 
leads to strides in the development of new technology.  But for the most part, I'll probably loosely 
follow the guide supplied by Google.  With that in mind, I'd like to mention a couple topics that 
I, along with many others, don't necessarily agree with. 

The first topic that I think should be mentioned is Google's take on exceptions.  I'm sure most 
other kids on this assignment are bringing up the same point, but it's definitely something that 
can't be ignored.  It was on a Quora post by Brian Bi where I found information about Google's 
take on exceptions \cite{brian}.  Essentially, their take is that you shouldn't use them because 
of their own code base, they say it would be a hassle to introduce exceptions to their existing 
exception-free code.  Now while they may be right that introducing exceptions now on their humungous 
project would be quite difficult, for anyone starting their own project, exceptions are a great way 
to debug errors in your program, and should definitely be adopted.

There are a couple other points the style guide goes over that I most likely won't use myself.  I'm 
not all that knowledgeable on these topics though, so this will be more on a matter of personal 
preference.  The style guide discourages multiple inheritence because of the possibility of 
complicated coad and overhead.  They may be right, I'm just a silly little Junior after all, but I know 
that the most basic C++ library (iostream) uses multiple implementation inheritence.  So while I'm not 
100\% sure I'll even reach a point where I might consider multiple inheritence, I definitely won't let 
the style guide influence my decision on this point.  The final point I'll be going over is the standard 
header guard.  The question on why we use the standard header guard as opposed to the #pragma was 
brought to my attention by a piazza post \cite{tom}.  I think it's an interesting choice to use the 
\texttt{\#pragma} method, and while I kinda doubt I'll be using it on my project, I do think it's a point 
that should have maybe been brought up in the style guide.  But with all this in mind, I think the style 
guide is actually very helpful and definitely gets a bad rap for little reason, but I'm definitely not 
treating it like some sort of holy text.  Stack Overflow is what I always refer to for good code practice, 
and I'll probably just stick to that, because it has many people to keep eachother in check.  In conclusion, 
I'll just do shit my own way and I apologize for writing too much again (and for swearing). \\\\

\bibliographystyle{apalike}
\bibliography{sources}
\end{document}
